---
title: "Submodels"
---

# *submodels function*

*This is the main function in the submodels family.*

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r, file="C:\\Users\\mbbx4sb5\\Dropbox (The University of Manchester)\\00_ICS_RECODE_Shared\\R\\prediction\\submodels.R", echo=TRUE, message=FALSE, warning=FALSE}
```

### Basic premise of the function

1.  Housekeeping and sanitizing inputs

2.  Getting the missingness patterns (see [Helper functions](helper_functions.qmd))

3.  Stops the computing of models that would have dropped the force variables.

4.  Set the threshold (uses the number of variables in the model)

# predict.submodels

```{r, file="C:\\Users\\mbbx4sb5\\Dropbox (The University of Manchester)\\00_ICS_RECODE_Shared\\R\\prediction\\predict_submodels.R", echo=TRUE, message=FALSE, warning=FALSE}
```

# auc

```{r, file="C:\\Users\\mbbx4sb5\\Dropbox (The University of Manchester)\\00_ICS_RECODE_Shared\\R\\prediction\\auc.R", echo=TRUE, message=FALSE, warning=FALSE}
```
